Supplement to ``{\mytitle}''

### 2022-10-14 ms.Rnw
Comments/cleaning up

<<ReplicateModel1, echo=FALSE, message=FALSE, include=FALSE>>=
# In this chunk, we replicate the results in Lee et al for model 1.
# We do this replication here in the ms.Rnw file because we need to get
# estimates for the likelihood of the models used by lee et al. The chunk has
# the following steps:
#   (1) Initialize some starting values close to those used in lee et al
#   (2) do small MIF, similar to what was done in lee et al, to get a fit for
#       epidemic period.
#   (3) Simulate from model
#   (4) Use simulations to initialize endemic phase of the model, and
#       perform MIF, similar to what lee et al did for this phase
#   (5) Simulate from model
#   (6) Calculate the likelihood of each possible set of parameters.

N_PARAMS   <- switch(RUN_LEVEL,  10,  20,  300)
N_LEE_SIMS <- switch(RUN_LEVEL,  10,  20,   20)
NMIF       <- switch(RUN_LEVEL,   5,  20,   50)
NP_MIF     <- switch(RUN_LEVEL,  20,  50,  100)  # Too few, but matches Lee et al
NP_EVAL    <- switch(RUN_LEVEL,  50, 200, 2000)
NREPS_EVAL <- switch(RUN_LEVEL,   3,   5,   36)

#-------------------------- Step 1 ----------------------

@@

#-------------------------- Step 2 ----------------------

@@

epi_params$sig_sq <- 0

# simply re-ordering to match model order
epi_params <- epi_params[, names(coef(h1_epi))]

epi_rw <- rw.sd(
  # Parameters close to desired values
  beta1 = 0.001,
  tau   = 0.001,
  nu    = 0.001,
  rho   = 0.001,
  # Parameters not close to desired values
  beta2 = 0.02,
  beta3 = 0.02,
  beta4 = 0.02,
  beta5 = 0.02,
  beta6 = 0.02,
  E_0   = ivp(0.1),
  I_0   = ivp(0.1)
)

@@

rm(epi_rw, epi_params, not_in_start_params,
   lower, upper, betas, bounds, starting_params, i)
gc()

# epi_complete <- as.data.frame(t(coef(lee1_epi_mif)))
# epi_complete$logLik <- logLik(lee1_epi_mif)

# GGally::ggpairs(
#   filter(epi_results, logLik > -2200),
#   columns = c("rho", "tau", "beta1", "nu", "logLik")
# )

# epi_complete <- filter(epi_complete, logLik > -2200)

if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_epi_evals.rds")) {
  lee_epi_ll <- readRDS("model1/run_level_3/lee1_epi_evals.rds")
} else {
  lee_epi_ll <- bake(
    file = paste0("model1/", rl_dir, "lee1_epi_evals.rds"), {

@@

#-------------------------- Step 3 ----------------------

# quants <- all_sims %>%
#   select(.id, week, cases) %>%
#   group_by(week) %>%
#   summarize(
#     q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
#     q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
#     q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
#   ) %>%
#   ungroup() %>%
#   mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

#-------------------------- Step 4 ----------------------

h1_end <- haiti1(period = "endemic")

end_rw <- rw.sd(
  beta1 = 0.05,
  tau   = 0.005,
  nu    = 0.04,
  rho   = 0.01,
  beta2 = 0.05,
  beta3 = 0.05,
  beta4 = 0.05,
  beta5 = 0.05,
  beta6 = 0.05
)

registerDoParallel(20)
registerDoRNG(21489587)

end_start <- lee1_epi_mif$lee1_epi_mif_params

# end_start <- epi_complete %>%
#   select(-logLik)

if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_end_fit.rds")) {
  lee1_end_mif <- readRDS("model1/run_level_3/lee1_end_fit.rds")
} else {
  lee1_end_mif <- bake(
    file = paste0("model1/", rl_dir, "lee1_end_fit.rds"),
    {
      foreach(
        i = 1:nrow(end_start),
        .packages = c('pomp'),
        .combine = rbind
      ) %dopar% {

        output <- list()
        r_params <- end_start[i, ]

        lee1_epi_mif$start_states %>%
          filter(param_set == i) %>%
          select(-param_set) %>%
          sapply(., median, na.rm = TRUE) -> rinit_parms

        # Set the initial state
        new_rinit <- Csnippet(
          sprintf(
            "
          S = nearbyint(%f);
          E = nearbyint(%f);
          I = nearbyint(%f);
          A = nearbyint(%f);
          R = nearbyint(%f);
          incid = nearbyint(%f);
          foival = %f;
          Str0 = nearbyint(%f);
          Sout = nearbyint(%f);
          Sin = nearbyint(%f);
          ",
          rinit_parms['S'],
          rinit_parms['E'],
          rinit_parms['I'],
          rinit_parms['A'],
          rinit_parms['R'],
          rinit_parms['incid'],
          rinit_parms['foival'],
          rinit_parms['Str0'],
          rinit_parms['Sout'],
          rinit_parms['Sin']
          )
        )

        h1_end <- h1_end %>% pomp(
          rinit = new_rinit,
          statenames = c(
            "S", "E", "I", "A", "R", "incid",
            "foival", "Str0", "Sout", "Sin"
          )
        )

        coef(h1_end) <- r_params
        mf_end <- mif2(
          h1_end,
          Np = NP_MIF,
          Nmif = NMIF,
          cooling.type = "hyperbolic",
          cooling.fraction.50 = 0.05,  # Cooling set so that we will reach smaller rw.sd at the end than the end of the unit3 search.
          rw.sd = end_rw
        )

>>>>

<<CalculateLee1Likelihoods, echo=FALSE, message=FALSE, include=FALSE>>=

# In this code chunk, we attempt to estimate the likelihood of Model 1 that 
# is described by Lee et al (2020). Because this chunk produces some 
# computationally intensive calculations that must also be done in the 
# supplement, we save some additional information in this chunk that is only 
# used in the supplement (si/si.Rnw) and not in the manuscript.

# Setting algorithmic hyper-parameters based on RUN_LEVEL. 
# 
#                    RUN_LEVEL =  1,   2,    3
N_PARAMS   <- switch(RUN_LEVEL,  10,  20,  300)  # Number of parameters to sample
N_LEE_SIMS <- switch(RUN_LEVEL,  10,  20,   20)  # Number of simulations per parameter
NMIF       <- switch(RUN_LEVEL,   5,  20,   50)  # Number of MIF iterations
NP_MIF     <- switch(RUN_LEVEL,  20,  50,  100)  # Number of particles for IF2. This is too few for proper optimization, but matches Lee et al
NP_EVAL    <- switch(RUN_LEVEL,  50, 200, 2000)  # Number of particles used to evaluate the likelihood 
NREPS_EVAL <- switch(RUN_LEVEL,   3,   5,   36)  # Number of replicates of likelihood evaluation.

@@

@@

# The process model of Lee et al (2020) didn't include over-dispersion
epi_params$sig_sq <- 0

# re-ordering starting parameters to match model order
epi_params <- epi_params[, names(coef(h1_epi))]

# Setting the random walk standard deviation. 
epi_rw <- rw.sd(
  # Parameters close to desired values
  beta1 = 0.001,
  tau   = 0.001,
  nu    = 0.001,
  rho   = 0.001,
  # Parameters not close to desired values
  beta2 = 0.02,
  beta3 = 0.02,
  beta4 = 0.02,
  beta5 = 0.02,
  beta6 = 0.02,
  E_0   = ivp(0.1),
  I_0   = ivp(0.1)
)

@@

# Some clean-up of unused variables, to avoid "Environment clutter".
rm(epi_rw, epi_params, not_in_start_params,
   lower, upper, betas, bounds, starting_params, i)
gc()

# Check if the run_level_3 results exist. If they do, load these results. 
# If not, compute the run_level_<RUN_LEVEL> results and save them. 
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_epi_evals.rds")) {
  lee_epi_ll <- readRDS("model1/run_level_3/lee1_epi_evals.rds")
} else {
  lee_epi_ll <- bake(
    file = paste0("model1/", rl_dir, "lee1_epi_evals.rds"), {

@@

# Create the endemic version of the model
h1_end <- haiti1(period = "endemic")

# Our parameter values are now very similar to those used by Lee et al (2020). 
# We therefore use the same rw.sd that was used in their source code. 
end_rw <- rw.sd(
  beta1 = 0.05,
  tau   = 0.005,
  nu    = 0.04,
  rho   = 0.01,
  beta2 = 0.05,
  beta3 = 0.05,
  beta4 = 0.05,
  beta5 = 0.05,
  beta6 = 0.05
)

registerDoRNG(21489587)  # For reproducible computing

# Store the parameter values from the end of the IF2 search for the epidemic 
# model as the start values for the IF2 search for the endemic model, as was 
# described by Lee et al (2020)
end_start <- lee1_epi_mif$lee1_epi_mif_params

# Check if the run_level_3 results exist. If they do, load these results. 
# If not, compute the run_level_<RUN_LEVEL> results and save them. 
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_end_fit.rds")) {
  lee1_end_mif <- readRDS("model1/run_level_3/lee1_end_fit.rds")
} else {
  lee1_end_mif <- bake(
    file = paste0("model1/", rl_dir, "lee1_end_fit.rds"),
    {
      foreach(  # Parallel execution of the IF2 algorith. 
        i = 1:nrow(end_start),
        .packages = c('pomp'),
        .combine = rbind
      ) %dopar% {

        # Get the ith set of starting parameters for the endemic model. These 
        # correspond to the parameters resulting from the ith execution of the 
        # IF2 algorithm on the epidemic model
        r_params <- end_start[i, ]

        # Calculate the starting states for each set of IF2 parameters. Because 
        # there are <N_LEE_SIMS> simulations for each set of parameters, we 
        # take the median value for each set of parameters as the starting 
        # value for the endemic model. 
        lee1_epi_mif$start_states %>%
          filter(param_set == i) %>%
          select(-param_set) %>%
          sapply(., median, na.rm = TRUE) -> rinit_parms

        # Set the initial state by creating a new rinit function for endemic 
        # model. 
        new_rinit <- Csnippet(
          sprintf(
            "
          S = nearbyint(%f);
          E = nearbyint(%f);
          I = nearbyint(%f);
          A = nearbyint(%f);
          R = nearbyint(%f);
          incid = nearbyint(%f);
          foival = %f;
          Str0 = nearbyint(%f);
          Sout = nearbyint(%f);
          Sin = nearbyint(%f);
          ",
          rinit_parms['S'],
          rinit_parms['E'],
          rinit_parms['I'],
          rinit_parms['A'],
          rinit_parms['R'],
          rinit_parms['incid'],
          rinit_parms['foival'],
          rinit_parms['Str0'],
          rinit_parms['Sout'],
          rinit_parms['Sin']
          )
        )

        # Create a new pomp object representing the endemic model, where the 
        # starting state is the median of the latent states from the epidemic 
        # model for the parameter set i. 
        h1_end <- h1_end %>% pomp(
          rinit = new_rinit,
          statenames = c(
            "S", "E", "I", "A", "R", "incid",
            "foival", "Str0", "Sout", "Sin"
          )
        )

        # Set model coefficients to be the parameter set i.
        coef(h1_end) <- r_params
        mf_end <- mif2(
          h1_end,
          Np = NP_MIF,
          Nmif = NMIF,
          cooling.type = "hyperbolic",
          cooling.fraction.50 = 0.05,  # Matches Lee et al (2020)
          rw.sd = end_rw
        )

### 2022-10-14 si/si.Rnw
Comments/cleaning up

There are minor differences between Figure~\ref{fig:mod2rep} and Figure~S15 of \cite{lee20sup}.
% such as a delayed second wave in the epidemic period and a larger initial wave during the endemic period.
While the discrepancy appears minor, the deterministic nature of Model~2 implies that an exact replication of model trajectories should be possible.
In this case, these discrepancies may be attributed to model implementation and small differences in model parameters.
Details about the measurement models and how latent states were initialized for the epidemic model were not provided by \citet{lee20sup} and therefore these details must be inferred by looking at the provided source code.
According to repository comments, the files \code{fitInPieces3paramsCleanMay2019Public.py} and \code{fitInPiecesMuWithFracSusFixedAllInfectionsPublic.py} were used to fit the epidemic and endemic phases of the model respectively, although it is apparent that these exact files were not used to obtain the reported results.
These files contain errors that making them impossible to run, and therefore these files cannot be the code that generated the reported model fits\footnote{One example of why the code cannot be run that the file loads functions from a non-extant file named \code{choleraEqs.py} in line 13 rather than \code{choleraEqsPublic.py}.}.
The inability to replicate the results by \citet{lee20} by running the provided source code makes checking whether or not a our numeric implementation faithfully represents their results very difficult.
% The difficulty of reproducing the result of \citet{lee20} is coupled with a misunderstanding of how these results were originally obtained.

\jwc{PROBABLY DELETE THIS NEXT BIT, BUT IT'S USEFUL TO DOCUMENT IT FOR NOW...}
One concern that arises when inspecting the files \code{fitInPieces3paramsCleanMay2019Public.py} and \code{fitInPiecesMuWithFracSusFixedAllInfectionsPublic.py} is that the output of these files (which contains the estimated parameters) was not referenced/used in either the file that generates the sensitivity analysis or the file that generates the model simulations; instead, alternative (non-extant) files with similar names are loaded, suggesting parameters that were used to generate the final results were not a direct result of the estimated parameters.
Furthermore, inspection of the script \code{simulateHaitiFullTimeInPiecesAllVaccinationScenariosSingleParameterVectorAllInfectionsPublic.py}---which is used to obtain simulations for each of the vaccination scenarios---results in a number of concerns.
First, the script uses parameters from the pickle file \code{leastSquaresDataUpTo2019/fitEpidemicPiece\_resultsMatrix\_18Mar2019.pickle}, which only appears in the repository as a commented out line of code in the file \code{generateSetsOfParametersForSensitivityAnalysisPublic.py}, suggesting that the set of parameters used to simulate each vaccination scenario was not the most recent set of parameters generated by the sensitivity analysis.
Second, the script appears to use the set of differential equations from the function \code{choleraEqs11WithVaccinationNetwork} for the epidemic phase, even though parameter values for this phase were estimated using \code{choleraEqs10WithVaccinationNetwork}.
It is possible, however, that we simply are interpreting the code incorrectly and that these concerns are unwarranted.
Furthermore, because the provided code is non-functional, it is possible that the code that was actually used to obtain the reported results does not include these potential errors.
This highlights the importance of providing the exact source code that was used to obtain the reported results of any analysis.

>>>>

There are minor differences between Figure~\ref{fig:mod2rep} and Figure~S15 of \cite{lee20sup}.
% such as a delayed second wave in the epidemic period and a larger initial wave during the endemic period.
While the discrepancy appears minor, the deterministic nature of Model~2 implies that an exact replication of model trajectories should be possible.
In this case, these discrepancies may be attributed to model implementation and small differences in model parameters.
Details about the measurement models and how latent states were initialized for the epidemic model were not provided by \citet{lee20sup} and therefore these details must be inferred by looking at the provided source code.
According to repository comments, the files \code{fit\allowbreak In\allowbreak Pieces\allowbreak 3params\allowbreak Clean\allowbreak May2019\allowbreak Public.py} and \code{fit\allowbreak In\allowbreak Pieces\allowbreak Mu\allowbreak With\allowbreak Frac\allowbreak Sus\allowbreak Fixed\allowbreak All\allowbreak Infections\allowbreak Public.py} were used to fit the epidemic and endemic phases of the model respectively, although it is apparent that these exact files were not used to obtain the reported results.
These files contain errors that making them impossible to run, and therefore these files cannot be the code that generated the reported model fits\footnote{One example of why the code cannot be run that the file loads functions from a non-extant file named \code{choleraEqs.py} in line 13 rather than \code{cholera\allowbreak Eqs\allowbreak Public.py}.}.
The inability to replicate the results by \citet{lee20} by running the provided source code makes checking whether or not a our numeric implementation faithfully represents their results very difficult.
% The difficulty of reproducing the result of \citet{lee20} is coupled with a misunderstanding of how these results were originally obtained.

\jwc{PROBABLY DELETE THIS NEXT BIT, BUT IT'S USEFUL TO DOCUMENT IT FOR NOW...}
One concern that arises when inspecting the files \code{fit\allowbreak In\allowbreak Pieces3\allowbreak params\allowbreak Clean\allowbreak May\allowbreak 2019\allowbreak Public.py} and \code{fit\allowbreak In\allowbreak Pieces\allowbreak Mu\allowbreak With\allowbreak Frac\allowbreak Sus\allowbreak Fixed\allowbreak All\allowbreak Infections\allowbreak Public.py} is that the output of these files (which contains the estimated parameters) was not referenced/used in either the file that generates the sensitivity analysis or the file that generates the model simulations; instead, alternative (non-extant) files with similar names are loaded, suggesting parameters that were used to generate the final results were not a direct result of the estimated parameters.
Furthermore, inspection of the script \code{simulate\allowbreak Haiti\allowbreak Full\allowbreak Time\allowbreak In\allowbreak Pieces\allowbreak All\allowbreak Vaccination\allowbreak Scenarios\allowbreak Single\allowbreak Parameter\allowbreak Vector\allowbreak All\allowbreak Infections\allowbreak Public.py}---which is used to obtain simulations for each of the vaccination scenarios---results in a number of concerns.
First, the script uses parameters from the pickle file \code{least\allowbreak Squares\allowbreak Data\allowbreak UpTo\allowbreak 2019\allowbreak /fit\allowbreak Epidemic\allowbreak Piece\_results\allowbreak Matrix\_18\allowbreak Mar\allowbreak 2019.pickle}, which only appears in the repository as a commented out line of code in the file \code{generate\allowbreak Sets\allowbreak Of\allowbreak Parameters\allowbreak For\allowbreak Sensitivity\allowbreak Analysis\allowbreak Public.py}, suggesting that the set of parameters used to simulate each vaccination scenario was not the most recent set of parameters generated by the sensitivity analysis.
Second, the script appears to use the set of differential equations from the function \code{cholera\allowbreak Eqs11\allowbreak With\allowbreak Vaccination\allowbreak Network} for the epidemic phase, even though parameter values for this phase were estimated using \code{cholera\allowbreak Eqs10\allowbreak With\allowbreak Vaccination\allowbreak Network}.
It is possible, however, that we simply are interpreting the code incorrectly and that these concerns are unwarranted.
Furthermore, because the provided code is non-functional, it is possible that the code that was actually used to obtain the reported results does not include these potential errors.
This highlights the importance of providing the exact source code that was used to obtain the reported results of any analysis.
